---
title: "Homework 9 - SDS315"
author: "Alaina Gomez (agg3395)"
date: "2024-04-14"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
  pdf_document:
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo = FALSE, include = FALSE}
library(tidyverse)
library(mosaic)
library(ggplot2)
library(dplyr)
library(MatchIt)
library(tableone)
gotv <- read.csv("/Users/alainagomez/Desktop/turnout.csv")
solder <- read.csv("/Users/alainagomez/Desktop/solder.csv")
```

# Problem 1

## Part A

```{r, echo = FALSE}
prop_a <- mean(gotv$GOTV_call == 1 & gotv$voted1998 == 1)
prop_b <- mean(gotv$GOTV_call == 0 & gotv$voted1998 == 1)

prop_c <- prop_a - prop_b

# simulate a different proportion
prop_function <- function(gotv) {
  prop_1 <- mean(gotv$GOTV_call == 1 & gotv$voted1998 == 1)
  prop_2 <- mean(gotv$GOTV_call == 0 & gotv$voted1998 == 1)
  return(prop_1 - prop_2)
}

# bootstrap
unmatch_1998 <- do(10000)*prop_function(resample(gotv))
confint(unmatch_1998, level = 0.95)
```

The proportion of those receiving a GOTV call who voted in 1998 is
`r round(prop_a, 3)`.

The sample proportion of those not receiving a GOTV call who voted in
1998 is `r round(prop_b, 3)`.

The large sample 95% confidence interval for the difference in these two proportions is in the table above.

## Part B

```{r, echo = FALSE}
yes_call <- filter(gotv, GOTV_call == 1)
no_call <- filter(gotv, GOTV_call == 0)

age_diff <- mean(yes_call$AGE) - mean(no_call$AGE)
vote_diff <- mean(yes_call$voted1996) - mean(no_call$voted1996)
party_diff <- mean(yes_call$MAJORPTY) - mean(no_call$MAJORPTY)

confound_table <- data.frame(
  Variable = c("voted1996", "AGE", "MAJORPTY"),
  Prop_Diff_Call = c(vote_diff, age_diff, party_diff)
)
print(confound_table)

basic <- lm(voted1998 ~ GOTV_call, data = gotv)
confound_model <- lm(voted1998 ~ GOTV_call + voted1996 + AGE + MAJORPTY, data = gotv)
```

This table represents the difference in proportions of different
variables of those who received a GOTV call vs. those who did not.

```{r, echo = FALSE}
summary(basic)
```

```{r, echo = FALSE}
summary(confound_model)
```

## Part C

```{r, echo = FALSE}
matched <- matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, data = gotv, method = "nearest", ratio = 5)
match_data <- match.data(matched)

balance_before <- CreateTableOne(vars = c("voted1996", "AGE", "MAJORPTY", "voted1998"), data = gotv)
balance_after <- CreateTableOne(vars = c("voted1996", "AGE", "MAJORPTY", "voted1998"), strata = "GOTV_call", data = match_data)

print(balance_before, smd = TRUE)
print(balance_after, smd = TRUE)
```

While there is slight variation in the voted1996, AGE, and MAJORPTY variables, the p-values between treatment and control groups are substantially greater than 0.05. Matching has successfully balanced these covariates between the groups. Any information on "voted1998" should be ignored, only included for later use.

```{r, echo = FALSE}
prop_e <- mean(match_data$GOTV_call == 1 & match_data$voted1998 == 1)
prop_f <- mean(match_data$GOTV_call == 0 & match_data$voted1998 == 1)

prop_g <- prop_e - prop_f

# bootstrap
match_1998 <- do(10000)*prop_function(resample(match_data))
confint(match_1998, level = 0.95)
```

The proportion of those receiving a GOTV call who voted in 1998 is
`r round(prop_e, 3)`.

The sample proportion of those not receiving a GOTV call who voted in
1998 is `r round(prop_f, 3)`.

There is a statistically significant difference in the likelihood of
voting in the 1998 election between those who received a GOTV call and
those who did not. This suggests that the GOTV call had a significant
effect on voter turnout in the 1998 election. It can be concluded that
**the GOTV call likely had a substantial negative effect on voter turnout**
as both the unmatched and matched confidence intervals are both negative
and do not include zero. Overall, this difference could have significant
implications for electoral outcomes and campaign strategies.

# Problem 2

## Part A

```{r, echo = FALSE}
plot1 <- ggplot(solder, aes(x = Opening, y = skips)) + geom_boxplot(fill = "skyblue", color = "black") + labs(title = "Opening Size and Number of Skips", x = "Size of Opening of Solder Gun", y = "Number of Skips") + theme_minimal()

# Plot 2: Thickness of alloy vs. Number of skips
plot2 <- ggplot(solder, aes(x = Solder, y = skips)) + geom_boxplot(fill = "pink", color = "black") + labs(title = "Solder Thickness and Number of Skips", x = "Thickness of Alloy Used for Soldering", y = "Number of Skips") + theme_minimal()

# Display the plots
print(plot1)
print(plot2)
```

## Part B

```{r, echo = FALSE}
model <- lm(skips ~ Opening + Solder + Opening:Solder, solder)
coefficients_table <- summary(model)$coefficients
conf_intervals <- confint.default(model)
results <- cbind(coefficients_table, conf_intervals)

print(results)
```

## Part C

The intercept of the linear model is 0.393, which is only needed to represent the model.The main effect of the OpeningM is 2.407 skips. This is the effect of OpeningM in isolation. The main effect of the OpeningS is 5.127 skips. This is the effect of OpeningS in isolation. The main effect of SoldierThin is 2.280 skips. This is the effect of SoldierThin in isolation. The interaction effect of OpeningM:SoldierThin is -0.740 skips. Guns that have both "OpeningM" and "SoldierThin" yield an average of -0.74 less skips than what you would expect from summing the individual “isolated” effects of the two variables. The interaction effect of OpeningS:SoldierThin is 9.653 skips. Guns that have both "OpeningS" and "SoldierThin" yield an average of 0.653 skips more than what you would expect from summing the individual “isolated” effects of the two variables.

## Part D

Based on this analysis, I would recommend **a combination of OpeningM and SoldierThin thickness to AT&T because the combined effect of the interacting predictors is associated with fewer skips.** When accounting for the individual variables in comparison to the interaction, there is a negative association with the interaction. Meaning, there is a minimal number of steps in the manufacturing process.